---
title: "k_calcs_watertemp_data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE}
library(RPostgreSQL)
library(readr)
library(magrittr)
library(xts)
library(dplyr)
library(ggplot2)
```

```{r dbconnect, include=FALSE}
password <- scan(".pgpass", what="")
db <- dbConnect(PostgreSQL(), 
                host = "sesync-postgis01.research.sesync.org",
                dbname = "choptank", 
                user = "palmergroup",
                password = password)
```


Temperature profile data aren't in the database yet, so read from files.

```{r}
read_hobo_mx2202 <- function(filename){
  readr::read_csv(filename, skip = 2) %>%
  rename(datetime = `Date Time - GMT -05:00`,
         temp_F = `Temp, (*F)`,
         intensity_lmft2 = `Intensity, (lum/ft²)`) %>%
  dplyr::select(datetime, temp_F, intensity_lmft2)  %>%
    mutate(temp_C = (temp_F - 32)*(5/9))
}
```

```{r}
files <- list.files("../Delmarva/data/hobos", full.names = TRUE)
dl_dates <- sapply(files, function(x) substr(x, 36, 45)) %>% 
  unname() %>% unique() %>% as.Date()

qbMID <- grep(pattern = "20398209", files, value = TRUE) %>%
  grep(pattern = as.character(max(dl_dates)), ., value = TRUE) %>%
  read_hobo_mx2202()

qbTOP <- grep(pattern = "20398211", files, value = TRUE) %>%
  grep(pattern = as.character(max(dl_dates)), ., value = TRUE) %>%
  read_hobo_mx2202()
# qbMID2 <- grep(pattern = "2609", files, value = TRUE) %>%
#   grep(pattern = as.character(max(dl_dates)), ., value = TRUE) %>%
#   read_hobo_mx2202()
```

Aggregate five minute QB wind data to hourly
```{r}
qbMID_xts <- xts(qbMID$temp_C, order.by = qbMID$datetime)
ep <- endpoints(qbMID_xts, on = "hour")
temp_agg <- period.apply(qbMID_xts, INDEX = ep, FUN = mean)
```

```{r}
temp_df <- temp_agg %>%
  as.data.frame() %>%
  tibble::rownames_to_column(var = "datetime") %>%
  rename(temp_C = V1) %>%
  mutate(datetime = as.POSIXct(datetime, tz = ""))
```

```{r}
# qbTOP_xts <- xts(qbTOP$temp_C, order.by = qbTOP$datetime)
# ep <- endpoints(qbTOP_xts, on = "hour")
# temp_agg <- period.apply(qbTOP_xts, INDEX = ep, FUN = mean)

# temp_agg %>%
temp_df$temp_C %>%
  xts(order.by = temp_df$datetime) %>%
  dygraph() %>%
  dyRangeSelector() %>%
  dyLegend() %>%
  dyOptions(strokeWidth = 1) %>%
  # dyHighlight(highlightSeriesOpts = list(strokeWidth = 3)) %>%
  dyOptions(useDataTimezone = TRUE) %>%
  dyRoller(rollPeriod = 5) %>% # plotted point represents average of specified number of timestamps
  dyAxis("y", label = "water temp (C)")
```

Plots

```{r}
temp_df %>% 
  mutate(date = lubridate::date(datetime)) %>%
  mutate(time_hr = lubridate::hour(datetime)) %>% 
  group_by(date) %>%
  mutate(date_max = max(temp_C, na.rm = TRUE), 
         date_min = min(temp_C, na.rm = TRUE),
         date_range = date_max - date_min) %>%
  # filter(date >= as.Date('2018-06-01') & date <= as.Date('2018-07-01')) %>%
  ggplot(aes(x = date, y = date_range)) +
  # geom_line() +
  geom_line(aes(y = date_max, col = "red")) +
  geom_line(aes(y = date_min, col = "blue")) +
  # facet_wrap(~ lubridate::month(date)) +
  # facet_wrap(~ lubridate::yday(date)) +
  # facet_wrap(~ lubridate::wday(date, TRUE)) +
  # xlim(c(0, 24)) +
  theme_bw()
```
Looks like sensor was out of water from 8/26 to 9/10

```{r}
temp_df %>% 
  mutate(date = lubridate::date(datetime)) %>%
  mutate(time_hr = lubridate::hour(datetime)) %>% 
  mutate(yday = lubridate::yday(datetime)) %>%
  # group_by(date) %>%
  # mutate(date_max = max(temp_C, na.rm = TRUE), 
  #        date_min = min(temp_C, na.rm = TRUE),
  #        date_range = date_max - date_min) %>%
  filter(date >= as.Date('2018-07-23') & date <= as.Date('2018-08-25')) %>%
  ggplot(aes(x = time_hr, y = temp_C, group = date)) +
  geom_line(aes(col = yday), size = 2) +
  # geom_line(aes(y = date_max, col = "red")) +
  # geom_line(aes(y = date_min, col = "blue")) +
  # facet_wrap(~ lubridate::month(date)) +
  # facet_wrap(~ lubridate::yday(date)) +
  # facet_wrap(~ lubridate::wday(date, TRUE)) +
  xlim(c(0, 24)) +
  scale_color_viridis_c() +
  theme_bw()
```

8/26 to 9/10

```{r}
pdf("temp_by_weeks_QB.pdf", width = 10, height = 12)
temp_df %>%
  mutate(date = lubridate::date(datetime)) %>%
  mutate(time_hr = lubridate::hour(datetime)) %>% 
  mutate(yday = lubridate::yday(datetime)) %>%
  # group_by(date) %>%
  # mutate(date_max = max(temp_C, na.rm = TRUE),
  #        date_min = min(temp_C, na.rm = TRUE),
  #        date_range = date_max - date_min) %>%
  filter(date <= as.Date('2018-08-26') | date >= as.Date('2018-09-11')) %>%
  # mutate(datetime = as.POSIXct(datetime)) %>% dplyr::select(-date) %>%
  ggplot(aes(x = datetime, y = temp_C)) +
  # geom_line(aes(col = yday), size = 2) +
  geom_line(col = "gray") +
  geom_point(size = 2, aes(colour = time_hr)) +
  # geom_line(data = wl_df, aes(x = datetime, y = wl_m)) +
  # geom_line(aes(y = date_max, col = "red")) +
  # geom_line(aes(y = date_min, col = "blue")) +
  # facet_wrap(~ lubridate::month(date)) +
  # facet_wrap(~ lubridate::yday(date)) +
  # facet_wrap(~ lubridate::wday(date, TRUE)) +
  facet_wrap(~ lubridate::week(date), scales = "free_x", ncol = 2) +
  # xlim(c(0, 24)) +
  # scale_color_gradient2(low = "yellow", mid = "red", high = "blue", midpoint = 14) +
  scale_color_viridis_c() +
  theme_bw()
dev.off()
```
Typically, daily temperature peak around 2pm. 

```{r}
temp_df %>%
  mutate(date = lubridate::date(datetime)) %>%
  mutate(time_hr = lubridate::hour(datetime)) %>% 
  mutate(yday = lubridate::yday(datetime)) %>%
  # group_by(date) %>%
  # mutate(date_max = max(temp_C, na.rm = TRUE),
  #        date_min = min(temp_C, na.rm = TRUE),
  #        date_range = date_max - date_min) %>%
  filter(date >= as.Date('2018-08-01') & date <= as.Date('2018-08-12')) %>%
  # mutate(datetime = as.POSIXct(datetime)) %>% dplyr::select(-date) %>%
  ggplot(aes(x = datetime, y = temp_C)) +
  # geom_line(aes(col = yday), size = 2) +
  geom_line(col = "gray") +
  geom_point(size = 2, aes(col = time_hr)) +
  # geom_line(aes(y = date_max, col = "red")) +
  # geom_line(aes(y = date_min, col = "blue")) +
  # facet_wrap(~ lubridate::month(date)) +
  # facet_wrap(~ lubridate::yday(date)) +
  # facet_wrap(~ lubridate::wday(date, TRUE)) +
  # xlim(c(0, 24)) +
  scale_color_viridis_c() +
  theme_bw()
```

```{r}
temp_df %>%
  mutate(date = lubridate::date(datetime)) %>%
  mutate(time_hr = lubridate::hour(datetime)) %>% 
  mutate(yday = lubridate::yday(datetime)) %>%
  # group_by(date) %>%
  # mutate(date_max = max(temp_C, na.rm = TRUE),
  #        date_min = min(temp_C, na.rm = TRUE),
  #        date_range = date_max - date_min) %>%
  filter(date >= as.Date('2018-08-01') & date <= as.Date('2018-08-12')) %>%
  # mutate(datetime = as.POSIXct(datetime)) %>% dplyr::select(-date) %>%
  ggplot(aes(x = time_hr, y = temp_C, group = date)) +
  # geom_line(aes(col = yday), size = 2) +
  geom_line(size = 2, aes(col = yday)) +
  # geom_line(aes(y = date_max, col = "red")) +
  # geom_line(aes(y = date_min, col = "blue")) +
  # facet_wrap(~ lubridate::month(date)) +
  # facet_wrap(~ lubridate::yday(date)) +
  # facet_wrap(~ lubridate::wday(date, TRUE)) +
  xlim(c(0, 24)) +
  scale_color_viridis_c() +
  theme_bw()
```
```{r}
temp_df %>% 
  mutate(date = lubridate::date(datetime)) %>%
  mutate(time_hr = lubridate::hour(datetime)) %>% 
  # filter(date >= as.Date('2018-06-01') & date <= as.Date('2018-07-01')) %>%
  ggplot(aes(x = time_hr, y = temp_C, group = date)) +
  geom_line(aes(col = date)) +
  # facet_wrap(~ lubridate::month(date)) +
  # facet_wrap(~ lubridate::yday(date)) +
  # facet_wrap(~ lubridate::wday(date, TRUE)) +
  xlim(c(0, 24)) +
  theme_bw()
```

```{r}
temp_df %>% 
  mutate(date = lubridate::date(datetime)) %>%
  mutate(time_hr = lubridate::hour(datetime)) %>% 
  # group_by(date) %>%
  # mutate(date_max = max(temp_C, na.rm = TRUE), 
  #        date_min = min(temp_C, na.rm = TRUE),
  #        date_range = date_max - date_min) %>%
  filter(date >= as.Date('2018-09-11')) %>%
  ggplot(aes(x = time_hr, y = temp_C, group = date)) +
  geom_line() +
  # geom_line(aes(y = date_max, col = "red")) +
  # geom_line(aes(y = date_min, col = "blue")) +
  # facet_wrap(~ lubridate::month(date)) +
  facet_wrap(~ lubridate::yday(date)) +
  # facet_wrap(~ lubridate::wday(date, TRUE)) +
  xlim(c(0, 24)) +
  theme_bw()
```

```{r}
# difference object
dT <- diff(qbMID_agg[,1])
# plot(dT)
# head(dT)
```

Add column for water depth (Right now assuming static at 1m)

```{r}
q_ts <- merge(qbMID_agg, dT) %>%
  as.data.frame() %>%
  dplyr::rename(temp_C = qbMID_agg) %>%
  mutate(waterdepth_m = 1)
# q_ts <- merge(qbBOT_agg, dT) %>% 
#   as.data.frame() %>% 
#   rename(temp_C = qbBOT_agg) %>%
#   mutate(waterdensity = 
#            purrr::map_dbl(temp_C, ~calc_water_density_kgm3(temp_C = .x))) %>%
#   mutate(cp = 
#            purrr::map_dbl(temp_C, ~calc_Cp_JkgK(temp_C = .x))) %>%
#   mutate(waterdepth_m = 1) %>%
#   mutate(q = (dT/3600) * waterdensity * cp * waterdepth_m)
```

Apply function to calculate q to time series

```{r}
q_ts$q_wm2 <- q_ts %>% pmap_dbl(calc_q_heatflux)
```


```{r}
q_xts <- merge(dT, q_ts$q_wm2)
q_xts <- merge(q_xts, q_ts$temp_C)
q_xts <- merge(q_xts, q_ts$waterdepth_m)
# head(q_xts)
# plot(q_xts[,2])
```

```{r}
# dygraph(q_xts[,"q_ts.q_wm2"]) %>%
#   dyRangeSelector() %>%
#   dyLegend() %>%
#  dyAxis("y", label = "Q") %>%
#   dyOptions(strokeWidth = 1)
```




interactive plot for inspecting time series 

```{r}
# wind_xts %>%
#   dygraph() %>%
#   dyRangeSelector() %>%
#   dyLegend() %>%
#   dyOptions(strokeWidth = 1) %>%
#   # dyHighlight(highlightSeriesOpts = list(strokeWidth = 3)) %>%
#   dyOptions(useDataTimezone = TRUE) %>%
#   dyRoller(rollPeriod = 5) %>% # plotted point represents average of specified number of timestamps
#   dyAxis("y", label = "wind speed (m/s)")
```

Need to filter out the times when the sensor is out of the water, but for now that will not overlap with period with all three data sources (water level only through Sept 1?)

```{r}
write_csv(temp_df, "data/watertemp_qbMID.csv")
```


